### 2023-01-13 draft.tex
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@71f6201b5796e81becc17c4b88b6c3b0bb13905a 🚀
Deploying to gh-pages from @ gottacatchenall/ms_false_negatives@b0addac00a346b8ccd6a34ff1c11d4b9c9736697 🚀

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort. We suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate how the realized false-negative rate of
interactions is related to the relative abundance of the species pool,
and use simulation to produce a null estimate of the false-negative rate
given total sampling effort (the total count of all individuals of all
species seen), and to introduce a method for resampling of model
predictions of interaction probability to account for observation error.
We then show that positive associations in co-occurrence data can
increase the realized number of false-negatives, and demonstrate these
positive associations are rampant in network datasets, and conclude by
recommending that the simulation of sampling effort and species
occurrence can and should be used to help design surveys of species
interaction diversity (Moore \& McCarthy 2016), and by advocating use of
null models like those presented here as a tool for both guiding design
of surveys of species interactions and for modeling detection error in
predictive models.

@@

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}{%
\section{How many observations of a non-interaction do we need to be
confident it's a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}}

@@

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\section{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

@@

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates of unique niche
model networks per unique number of observations \(n\). All analyses
presented here are done in Julia v1.8 (Bezanson \emph{et al.} 2015)
using both EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville
\emph{et al.} 2021) and are hosted at (GITHUB\_LINK\_TODO). Note that
the empirical data, for the reasons described above, very likely already
contains many false negatives, we'll revisit this issue in the final
section.

@@

\hypertarget{resampling-interaction-probabilities-to-account-for-uncertainty-in-observation-error}{%
\section{Resampling interaction probabilities to account for uncertainty
in observation
error}\label{resampling-interaction-probabilities-to-account-for-uncertainty-in-observation-error}}

@@

@@

\hypertarget{positive-associations-increase-the-false-negative-rate}{%
\section{Positive associations increase the false-negative
rate}\label{positive-associations-increase-the-false-negative-rate}}

>>>>

Here, we illustrate how our confidence that a pair of species never
interacts highly depends on sampling effort. We suggest that surveys of
species interactions can benefit from simulation modeling of the
sampling process. We demonstrate how the realized false-negative rate of
interactions is related to the relative abundance of the species pool,
and use simulation to produce a null estimate of the false-negative rate
given total sampling effort (the total count of all individuals of all
species seen), and to introduce a method for introducing uncertainty
into model predictions of interaction probability to account for
observation error. We then show that positive associations in
co-occurrence data can increase the realized number of false-negatives,
and demonstrate these positive associations are rampant in network
datasets, and conclude by recommending that the simulation of sampling
effort and species occurrence can and should be used to help design
surveys of species interaction diversity (Moore \& McCarthy 2016), and
by advocating use of null models like those presented here as a tool for
both guiding design of surveys of species interactions and for modeling
detection error in predictive models.

@@

\hypertarget{accounting-for-false-negatives-in-species-interactions}{%
\section{Accounting for false-negatives in species
interactions}\label{accounting-for-false-negatives-in-species-interactions}}

\hypertarget{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}{%
\subsection{How many observations of a non-interaction do we need to be
confident it's a true
negative?}\label{how-many-observations-of-a-non-interaction-do-we-need-to-be-confident-its-a-true-negative}}

@@

\hypertarget{false-negatives-as-a-product-of-relative-abundance}{%
\subsection{False-negatives as a product of relative
abundance}\label{false-negatives-as-a-product-of-relative-abundance}}

@@

where the MAP estimate of \((\mu, \phi)\) applied to Mangal data from
(MacDonald \emph{et al.} 2020) is \((\mu=0.086, \phi=24.3)\). All
simulations were done with 500 independent replicates of unique niche
model networks per unique number of observations \(n\). All analyses
presented here are done in Julia v1.8 (Bezanson \emph{et al.} 2015)
using both EcologicalNetworks.jl v0.5 and Mangal.jl v0.4 (Banville
\emph{et al.} 2021) and are hosted on
\href{https://github.com/gottacatchenall/ms_false_negatives/tree/main/src}{Github}).
Note that the empirical data, for the reasons described above, very
likely already contains many false negatives, we'll revisit this issue
in the final section.

@@

\hypertarget{including-observation-error-in-interaction-predictions}{%
\subsection{Including observation-error in interaction
predictions}\label{including-observation-error-in-interaction-predictions}}

@@

There is also an analytic way to represent this.

Normal approximation to binomial, as the total number of samples from
\(n_p\) particles approaches infinity, the sum of total successes across
all particles approaches a normal distribution with parameters

\[\mathcal{N}\bigg(n_p p_{ij}, \sqrt{n_p p_{ij}(1-p_{ij})}\bigg)\]

Corrected \(p_{ij}\) based on \(p_fn\) and \(p_tn\) as
\(p_{ij}^* = p_{ij}(1+p_{fp})(1-p_{fn})\)
\[\mathcal{N}\bigg(n_p p_{ij}^*, \sqrt{n_p p_{ij}^* (1-p_{ij}^*)}\bigg)\]

and then by normalizing back to \((0,1)\).

\[\mathcal{N}\bigg(p^*_{ij}, \sqrt{\frac{ p_{ij}(1-p_{ij})}{n_p}} \bigg)\]

We can then further truncate this, although in practice probability mass
outside (0,1) is exteremly loss.

Number of particles \(n_p\) as a uncertainty width. Nautral analogue to
number of observations of co-occurrence.

@@

\hypertarget{positive-associations-in-co-occurrence-increase-the-false-negative-rate}{%
\section{Positive associations in co-occurrence increase the
false-negative
rate}\label{positive-associations-in-co-occurrence-increase-the-false-negative-rate}}
